Community-conditioned language models for linguistic variation

### 2021-12-12 paper/paper1.tex

Linguistic communication requires that speakers share
certain linguistic conventions, such as syntactic
structure, word meanings, and patterns of interaction.
Speakers assume that these conventions are \emph{common ground} among
their interlocutors, based on joint membership in a community
\cite{Stalnaker2002, Clark1996}.  Such a \emph{speech community} \citep{Gumperz1972} 
may be of any size, from the very small, like members of a particular friend group, 
to the very large, like speakers of English; however as \citet{Eckert1992}
point out, it is \emph{communities of practice}---defined by mutual social engagement
in some common activity---that are the primary locus of linguistic variation.

% Not only is linguistic variation an object of inquiry in its own
% right, but it is also important to consider it in the development of
% robust, equitable language technology.  Previous work has shown that
% by taking demographic information into account, models can achieve
% better performance on various standard NLP tasks
% \citep{Hovy2015,Yang2017}. \jp{Move this to related work so we can get to the point quicker.}
Most previous computational work on linguistic variation 
has considered variation at the level of macro-social categories, such as gender
\citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
geographic location \citep{Eisenstein2010,Bamman2014a}.
Recent exceptions are \citet{DelTredici2017} who use a modified
skip-gram model to show that lexical semantic variation occurs even across 
different communities organised around the same topic and 
\citet{Lucy2021}, who use contextualised word representations to
investigate community-level variation in word sense distributions.

In the present work, we investigate linguistic variation
in online communities on the social media website Reddit. 
In contrast to previous work, we investigate variation across 
speech communities, rather than macro-social categories.

@@

For this purpose, we introduce (\cref{sec:cclm}) various
Community-Conditioned Language Models (CCLMs for short). These models
are conditioned on a vector representation (or embedding), which
varies for each community. Hence, they learn \emph{community
  embeddings}.  Even though we test which architectures best make use
of the community information, our purpose is not to improve models in
terms of perplexity. Rather we want to extract the best set of
community embeddings and test in what ways the resulting embeddings
predict the social structure of subreddits. For this purpose, we test
the correlation with a social network-based representation of
communities (\cref{sec:embedding-analysis}).

@@

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}. Even though the transformer is dominating LSTM in
the state of the art, it achieves its dominance partly thanks to the
availability of very large data sets \jp{Bill: can you add relevant citations?
  (BERT, GPT3, etc.)},
which are not available to us.\footnote{Fine-tuning existing models
  falls outside our methodology, which consists in fundamentally
  changing the structure of the network by concatenating
  community embeddings.}.

In either case, the model is organised as a
standard $3$-layer neural sequence encoder, where the input for the
$t$th timestep of the $n+1$st layer is the $t$th hidden state of the 
$n$th layer.
As usual, the input to the first layer, is a sequence of tokens,
encoded with a trainable embedding layer over a pre-determined vocabulary.
At the other end, word tokens are predicted using a softmax projection 
layer. What we have described so far does not take community into account
and as such we call them \emph{uncodnitioned models}, but the same
encoder architecture also forms the core of our conditioned models.

In the conditional models, we add a \emph{community embedding} parameter.
This parameter is concatenated (at each time step) with the hidden layer of the
sequence encoder, at some layer $l_c \leq n$, and passed through a
linear layer which projects the resulting vector back to the original
hidden layer size.  For $l_c = n$, the output of this linear layer is
passed directly to the softmax function, just as the final hidden
layer of the sequence encoder is in other models.  For $l_c=0$, the
community embedding is concatenated with the token embedding.  For
this reason, we set the hidden size of the sequence encoder and the
size of the token embedding to be equal for all models.

%TODO: Diagram of model?

\subsection{Data sets}

We investigate linguistic variation across various communities 
from the social media website Reddit.\footnote{Comments were obtained
  from the archive at \url{https://pushshift.io/}.
  \cite{Baumgartner2020}.}
%
Reddit is divided into forums called \textit{subreddits}, 
which are typically organised around a topic of interest. 
Users create \textit{posts}, which consist of a link, image, 
or text, along with a \emph{comment} section. 
Comments are threaded: a comment can be made directly on a post,
or appear as a reply to another comment.
%
Hereafter we refer to such comments as ``messages'', matching our
convention in mathematical formulas: the letter $c$ stands for a
community, and $m$ stands for a message.

Our dataset includes messages from \num{510} subreddits, 
the set of all subreddits 
with at least \num{5000} messages per month for each month
of the year 2015.
Most but not all of these communities (with about 10 exceptions) 
are mostly English-language.
Each community corpus consist of \num{42000} randomly selected messages from the year 2015.
For each community, we reserve \num{1000} messages each for development and testing,
leaving a total of \num{20.4}M messages for training.

Messages were preprocessed as follows. 
We excluded the content of block quotes, code blocks, and tables.
We removed any markup (formatting) commands from the remaining text, 
extracting only rendered text.
We tokenized the messages using the default English model for the SpaCy tokenizer 
Version 2.2.3 \citep{Honnibal2017}.

@@

We use a vocabulary size of \num{40000} tokens, including a special
out-of-vocabulary token.  The vocabulary consists of the most frequent
tokens across all communities.

For the LSTM architecture, we trained the models on a simple left-to-right language
modeling task with cross-entropy loss.  Because
the transformer operates on all tokens in the sequence at once, the
language modeling was achieved by masking and incrementally un-masking
input tokens in a left-to-right fashion (so we have an
auto-regressive model).  We used the AdamW
\citep{Loshchilov2019} optimisation algorithm, with an initial
learning rate of \num{0.001}, with no extra control on the decay
of learning rate.
%
We used a batch size of \num{256} and a maximum sequence length of
\num{64} tokens, truncating longer messages (16.8\% of messages were
longer than \num{64} tokens).  During training, a dropout rate of
$0.1$ was applied between encoder layers and after each linear layer.

All experiments use models with \num{3} sequence encoder layers,
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen in part to give the LSTM and transformer
  models a comparable number of parameters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions. 
We experimented with every possible value for $l_c$ in a three-layer model ($l_c\in\{0,1,2,3\}$).

We trained the models until the validation loss stopped decreasing for
two epochs in a row, and used the weights from the epoch with the
smallest validation loss for testing.  In general, transformer models
were trained for about half as long as the LSTM models (\cref{tab:model-results}, 
test epoch).

@@

We also consider performance stratified by community; that is,
$\Ppl_{M_j}$ and $\IG_{M_j}$, 
where $M_j$ is the set of messages originating from community $c_j$ 
(\cref{fig:comm-stratified-box}).
We observe a lot of variation in baseline perplexity
across communities, with $\Ppl_{M_j}$ ranging from \num{3.67} to
\num{93.58} for the best conditional LSTM model 
(\cref{fig:comm-stratified-box}; also see supplementary materials for perplexity by community). %TODO keep? 
The conditioned models also perform differently across different communities---%
even among the best models, some communities have $\IG_{M_j} < 1$,
meaning that the CCLM performs worse than the unconditioned baseline for 
messages from that community.
For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
(\cref{fig:comm-stratified-box}).\footnote{
  Some of the communities with consistently high $IG_{M_j}$ 
  across all models are primarily non-English, 
  but surprisingly, not the three most extreme
  outliers. There are \subreddit{counting}, \subreddit{friendsafari}, and \subreddit{Fireteams}, 
  the later two of which are places where people coordinate to play video games
  together. The messages in these communites adhere to highly regular formats, 
  which are presumably conventional to the communities.}
%
%The variation in information gain across communities is not 
%related to perplexity, however (\cref{fig:ppl-info-gain}).

\newcommand{\modelboxplot}[3]{
  \addplot+[
    boxplot={draw position=#3, box extend=0.3}, 
    draw=#2, mark=*, mark options={fill=#2, scale=0.5}, solid, fill=#2!10,
    area legend] 
    table [y=#1] {floats/comm.csv};
}
\begin{figure}[t]
\begin{tikzpicture}
  \begin{axis}[name=ig,
    height=7cm,width=8cm,
    boxplot/draw direction=x, axis x line*=bottom, axis y line*=left,y axis line style={opacity=0},
    xmax=1.6,
    ylabel=$l_c$, xlabel=$\IG_{M_j}$,
    legend style={font=\tiny,at={(1,1)},anchor=north east}, legend columns=2
    ]
    \addplot[thin, red, samples=2] coordinates {(1,-0.5)(1,3.5)};
    \foreach \lc in {0, ..., 3}{
      \modelboxplot{lstm-3-\lc-ig}{violet}{\lc+0.2}
      \modelboxplot{transformer-3-\lc-ig}{teal}{\lc-0.2}
    \legend{,LSTM, Transformer}
    }
  \end{axis}
  \begin{axis}[name=ppl,
    anchor=north west,at={($(ig.south west)-(0,1.1cm)$)},
    width=8cm, height=2.5cm, 
    boxplot/draw direction=x, axis x line*=bottom, hide y axis,
    xlabel=$\Ppl_{M_j}$,
    ]
    \modelboxplot{lstm-3-ppl}{violet}{0.2}
    \modelboxplot{transformer-3-ppl}{teal}{-.2}
  \end{axis}
\end{tikzpicture}
\caption{
  Average model performance by community.
The boxes indicate upper and lower quartiles, while the whiskers are placed at the upper and lower maximum, with communities more than $1.5\times IQR$ (inter-quartile range) above the upper quartile considered outliers (represented as dots). The three most extreme outliers are excluded from this view.}.% \cref{fig:ppl-info-gain}).
   %See also table 1 in the supplementary materials % not very useful for this actually since we sort by indisc and it's only one model anyway.
   %for a tabular version of these results with community names.
\label{fig:comm-stratified-box}
\end{figure}

We observe empirically that across all the models we
tested, communities where conditioning has the least effect tend to be
organised around more general interest topics, such as 
\subreddit{relationships} and \subreddit{advice}, where the subject
matter is relevant to a broad range of people.  Conditioning the model
on community has the most benefit for narrower special-interest
subreddits, such as those organised around a certain videogame, sports
team, or subculture.  This makes sense intuitively, since communities
with more niche topics would tend to have more specialised vocabulary
and other linguistic patterns.

@@

In this section we investigate whether (and how much) community
embeddings obtained from the CCLMs correlate with the community-member
relationship on Reddit.

To this end, we compare the CCLM-learned community
embeddings\footnote{
  In this section, we only consider the embeddings from 
  the \emph{best} (highest information gain) CCLM from each architecture family; 
  that is, the LSTM with $l_c=1$ and the Transformer with $l_c=0$, however we
  observed similar results for other values of $l_c$.}
with the community embedding created by \citet{Kumar2018},
which were generated using using a negative-sampling optimization
algorithm, with the message author-community co-occurrence matrix as ground
truth, using data from January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only reflect the social
relationship between communities, via 
the user-community membership relation. In contrast, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%

@@

Unsuprisingly, the first category (\cref{fig:embedding-differences-examples}, left)yields communities that are qualitatively very similar. 
The \subreddit{SSBPM} and \subreddit{darksouls} communities are focused
around discussion of a particular videogame, and are paired
with communities that discuss a variation of the same game.
The \subreddit{amiugly} and \subreddit{Rateme} communities are both 
forums where the posts are selfies and the comments are mostly 
comments on the person's appearance.
The two communities paired with \subreddit{reddevils} are 
likewise comprised of fans of a particular English football club.
That the three football club communities are so closely related points
to an important point about the social embedding: Because of the way
it was trained, it does not have to be the case that communities with
a similar representation have a high overlap in their membership,
but rather that the \emph{kinds of members} of the first community
are also the \emph{kinds of members} who would belong to the second, 
based on their membership in other communities.

Communities with similar linguistic embeddings but dissimilar social embeddings
(\cref{fig:embedding-differences-examples}, left) tend to share a similar
topic, mode of interaction, or language variety, but 
in all cases we looked at, there is some reason to expect that
they might nevertheless attract different members.
For example,
\subreddit{hiphopheads} and \subreddit{Monstercat} are both topically
related to music, but the music genres are different, and the later
has a more geographically local focus (Monstercat is an independent
electronic music label based in Vancouver). The interactions in both 
\subreddit{MLPLounge} and \subreddit{CasualConversation} 
could be described as casual conversation, the former is intended specifically for
members of a niche internet sub-culture. % adult fans of the children's television show My Little Pony. There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
The \subreddit{exmormon} and \subreddit{Catholicism} communities discuss
the Mormon and Catholic churches, although their members have different 
relationships towards those organizations---the former is intended for
former members of the church, whereas the later is geared towards practicing
Catholics.
Finally, both \subreddit{rocketbeans} and \subreddit{de} are primarily
German-language subreddits, but the former is comprised of fans of a
computer gaming-related YouTube channel, while the later is more
general-interest.

@@

\subsection{Comparing embeddings: procrustes method}
\label{sec:statistical-evidence}

The anecdotal evidence suggests that for some communities, proximity
in linguistic patterns do match social proximity… and for others it
does not. Unfortunately, it seems hard, if not impossible, to draw a
general conclusion from this way. In this section, we propose a
systematic approach thanks to which we can quantify the correlation
between social proximity and linguistic proximity, and measure its
statistical significance.

@@

Here, the orthogonal matrix \(Ω\) gives a map from linguistic embeddings to social embeddings.
The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}.\footnote{
This approach has also been used to compare word embeddings across representations \citep[e.g.,][]{Hamilton2016a}.}
The solution is
\[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
decomposition (SVD) \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). That is, each
singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

As common when doing SVD, we can arrange \(U\), \(V\) and \(Σ\) such
that \(σ_i > σ_j\) iff \(i < j\). Doing so, the largest singular value
\(σ_0\) corresponds to the principal directions of correlation
(\(U_0\), \(V_0\)), \(σ_1\) to the second principal direction, etc.

@@

Thus if the real \(d(L,S)\) is below the mean by several standard
deviations, we can safely assume that there is statistically
significant correlation between \(L\) and \(S\). A 4-sigma difference
has less than one percent chance of occurring randomly. In our case:
we observe a difference between 61 and 68 standard deviations
(\cref{tab:embedding-correlations}). This definitely indicates a
significant correlation.  Furthermore, by coming back to the
definition of \(d(L,S)\), we know that, on average, the cosine
similarity between \(ΩL\) and \(S\) is \(0.45 = \) (\(510 - 232.18 / 510\)).  It
further means that if we obtain a linguistic embedding \(L_k\) for a
new community $k$, we can estimate its social embedding by \(ΩL_k\),
and the cosine similarity with its true social embedding \(S_k\) is
expected to be \(0.39\) = \((431.39 - 232.18) / 510\) --- accounting for
over-fitting effects by taking the average distance rather than the
maximum. % JP: TODO is the number from the table. this should probably be rewritten
In sum, it is clear that the CCLM embeddings predict some aspect the social-network embeddings--- but far from all of it.

@@

There are several recent studies that aim to measure \emph{linguistic distinctiveness}
at the level of speech community, which is an possible interpretation of 
of community-stratified information gain the CCLM over its unconditioned counterpart \cref{sec:ig}. Whereas the metrics in previous work are based on lexical frequency
\citep{OConnor2010,Zhang2018,Lucy2021}, CCLM information gain is capable of capturing 
distinctiveness at other levels of linguistic analysis. Further work is needed to
investigate exactly what kinds of variation are captured.

@@

The pattern of information gain by community is similar across
architectures.  Communities that benefit most from the conditioned
model behave that way for both the LSTM and transformer.  However,
there are some differences.  For example, many of the communities with the biggest 
difference in information gain between the $l_c=0$ and $l_c=3$ LSTMs 
seem to be organised around trading collectables or organising virtual meetups
(\emph{friendsafari},
\emph{Fireteams},
\emph{RandomActsOfGaming},
\emph{randomactsofcsgo},
\emph{Pokemongiveaway},
\emph{ACTrade},
\emph{SVExchange})
Why was the model where the community information was incorporated before
the LSTM able to take better advantage of it for these communities in particular? 
It would be interesting to investigate these differences in future work, since it
could reveal differences in the kind of linguistic variation the
different model architectures capture.

@@

Our main result is that the representation of communities built by
CCLM has a positive correlation with user co-occurrence patterns.
Even though such \emph{homophilic} correlation is a core hypothesis of 
sociolinguistics (see \citet{Kovacs2020}, for example), 
we believe that this study is the first to test it at 
the level of communities of practice using computatial methods.
Furthermore, it appears that our method (correlating
linguistic embeddings and social embeddings) is novel. Indeed, even
though the procrustes method has been used to correlate two sets of
linguistic embeddings \emph{for the same model}, we find no evidence
of the method being applied to embeddings for widely different models,
as we have done.

@@

\section{Ethical considerations}

>>>>

Linguistic communication requires that speakers share
certain linguistic conventions, such as syntactic
structure, word meanings, and patterns of interaction.
Speakers assume that these conventions are \emph{common ground} among
their interlocutors, based on joint membership in a community
\cite{Stalnaker2002, Clark1996}.  Such \emph{speech communities} \citep{Gumperz1972} 
range in size from the very small, like members of a friend group, 
to the very large, like speakers of English; however as \citet{Eckert1992}
point out, it is \emph{communities of practice}---defined by mutual social engagement
in a common activity---that are the primary locus of linguistic variation.

% Not only is linguistic variation an object of inquiry in its own
% right, but it is also important to consider it in the development of
% robust, equitable language technology.  Previous work has shown that
% by taking demographic information into account, models can achieve
% better performance on various standard NLP tasks
% \citep{Hovy2015,Yang2017}. \jp{Move this to related work so we can get to the point quicker.}
Variation is an important object of study in sociolinguistics,
and is naturally suited for computational analysis \citep{Nguyen2016}.
Most previous computational work on linguistic variation 
has considered variation at the level of macro-social categories, such as gender
\citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
geographic location \citep{Eisenstein2010,Bamman2014a}.%
\footnote{Recent exceptions are \citet{DelTredici2017} who use a modified
skip-gram model to show that lexical semantic variation occurs even across 
different communities organised around the same topic and 
\citet{Lucy2021}, who use contextualised word representations to
investigate community-level variation in word sense distributions.}
In the present work, we investigate linguistic variation
in online communities on the social media website Reddit. 
In contrast to previous work, we investigate variation across 
speech communities, rather than macro-social categories.

@@

For this purpose, we introduce (\cref{sec:cclm}) various
Community-Conditioned Language Models (CCLMs for short). These models
are conditioned on a vector representation (or embedding), which
varies for each community. Hence, they learn \emph{community
  embeddings}.  Although we test which architectures best make use
of the community information, our purpose is not to improve models in
terms of perplexity, but rather we want to extract the best set of
community embeddings and test how the resulting embeddings
predict the social structure of subreddits. For this purpose, we test
the correlation with a social network-based representation of
communities (\cref{sec:embedding-analysis}).

@@

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}. Although Transformer-based language models are considered
state of the art, they achieve dominance partly thanks to the
availability of very large data sets \citep[e.g.,][]{Devlin2019,Brown2020},
which are not available to us.\footnote{Fine-tuning existing models
  is not compatible with our methodology, since we would fundamentally
  change the structure of the network by concatenating
  community embeddings.}

In either case, the model is organised as a
standard $3$-layer neural sequence encoder, where the input for the
$t$th timestep of the $n+1$st layer is the $t$th hidden state of the 
$n$th layer.
As usual, the input to the first layer, is a sequence of tokens,
encoded with a trainable embedding layer over a pre-determined vocabulary.
At the other end, word tokens are predicted using a softmax projection 
layer. What we have described so far does not take community into account
and as such we call them \emph{unconditioned models}, but the same
encoder architecture also forms the core of our conditioned models.

In the CCLMs, we add a \emph{community embedding} parameter.
This parameter is concatenated (at each time step) with the hidden layer of the
sequence encoder, at some layer $l_c \leq n$, and passed through a
linear layer which projects the resulting vector back to the original
hidden layer size.  For $l_c = n$, the output of this linear layer is
passed directly to the softmax function, just as the final hidden
layer of the sequence encoder is in other models.  For $l_c=0$, the
community embedding is concatenated with the token embedding.  For
this reason, we set the hidden size of the sequence encoder and the
size of the token embedding to be equal for all models.

\subsection{Data sets}\label{sec:data}

We investigate linguistic variation across various communities 
from the social media website Reddit.\footnote{Comments were obtained
  from the archive at \url{https://pushshift.io/}.
  \cite{Baumgartner2020}. Code for reproducing our dataset, as well as
 our pre-trained community embeddings are available at URL.} %TODO ADD URL for camera ready
%
Reddit is divided into forums called \textit{subreddits}, 
which are typically organised around a topic of interest. 
Users create \textit{posts}, which consist of a link, image, 
or text, along with a \emph{comment} section. 
Comments are threaded: a comment can be made directly on a post,
or appear as a reply to another comment.
%
Hereafter we refer to such comments as ``messages'', matching our
convention in mathematical formulas: the letter $c$ stands for a
community, and $m$ stands for a message.

Our dataset includes messages from \num{510} subreddits, 
the set of all subreddits 
with at least \num{5000} messages per month for each month
of the year 2015.
Ignoring empty and deleted comments, we randomly sampled
\num{42000} messages from 2015 for each community.
We reserved \num{1000} messages from each community for development and testing,
leaving a total of \num{20.4}M messages for training.

Using \texttt{langid.py} \citep{Lui2012}, we observe that a majority of the overall
messages are classified as English (94.6\% of the test set) and X of \num{510} communities %TODO
have more than half of their messages classified as English.\footnote{See \cref{sec:ethics} for futher
  discussion.}

Messages were preprocessed as follows:
We excluded the content of block quotes, code blocks, and tables and
removed markup (formatting) commands, 
extracting only rendered text.
Messages were tokenized using the default English model for the SpaCy tokenizer 
Version 2.2.3 \citep{Honnibal2017}.

@@

The models use a vocabulary of \num{40000} tokens (including a special
out-of-vocabulary token), consisting of the most frequent tokens
across all communities.

We trained the models on a simple 
auto-regressive language modeling task with cross-entropy loss.  Because
the transformer operates on all tokens in the sequence at once, the
inputs to the model were masked and incrementally un-masked.
We used the AdamW
\citep{Loshchilov2019} optimisation algorithm, with an initial
learning rate of \num{0.001}, with no extra control on the decay
of learning rate.
%
The batch size was \num{256} and the maximum sequence length set to
\num{64} tokens, truncating longer messages (16.8\% of messages were
longer than \num{64} tokens).  During training, a dropout rate of
$0.1$ was applied between encoder layers and after each linear layer.

All experiments use models with \num{3} encoder layers,
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen to give the LSTM and transformer
  models a comparable number of parameters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions. 
We experimented with every possible value for $l_c$ in a three-layer model ($l_c\in\{0,1,2,3\}$).

We trained the models until the validation loss stopped decreasing for
two epochs in a row, and used the weights from the epoch with the
smallest validation loss for testing.

%In general, transformer models
%were trained for about half as long as the LSTM models (\cref{tab:model-results}, 
%test epoch).

@@

We also consider performance stratified by community; that is,
$\Ppl_{M_j}$ and $\IG_{M_j}$, 
where $M_j$ is the set of messages originating from community $c_j$ 
(\cref{fig:comm-stratified-box}).
We observe a lot of variation in baseline perplexity
across communities, with $\Ppl_{M_j}$ ranging from \num{3.67} to
\num{93.58} for the best conditional LSTM model 
(\cref{fig:comm-stratified-box}; also see supplementary materials for perplexity by community). %TODO keep? 
The conditioned models also perform differently across different communities---%
even among the best models, some communities have $\IG_{M_j} < 1$,
meaning that the CCLM performs worse than the unconditioned baseline for 
messages from that community.
For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
(\cref{fig:comm-stratified-box}).\footnote{
  Some of the communities with consistently high $IG_{M_j}$ 
  across all models are primarily non-English, 
  but surprisingly, not the three most extreme
  outliers. There are \subreddit{counting}, \subreddit{friendsafari}, and \subreddit{Fireteams}, 
  the later two of which are places where people coordinate to play video games
  together. The messages in these communites adhere to highly regular formats, 
  which are presumably conventional to the community.}
%
%The variation in information gain across communities is not 
%related to perplexity, however (\cref{fig:ppl-info-gain}).

\newcommand{\modelboxplot}[3]{
  \addplot+[
    boxplot={draw position=#3, box extend=0.3}, 
    draw=#2, mark=*, mark options={fill=#2, scale=0.5}, solid, fill=#2!10,
    area legend] 
    table [y=#1] {floats/comm.csv};
}
\begin{figure}[t]
\begin{tikzpicture}
  \begin{axis}[name=ig,
    height=7cm,width=8cm,
    boxplot/draw direction=x, axis x line*=bottom, axis y line*=left,y axis line style={opacity=0},
    xmax=1.6,
    ylabel=$l_c$, xlabel=$\IG_{M_j}$,
    legend style={font=\tiny,at={(1,1)},anchor=north east}, legend columns=2
    ]
    \addplot[thin, red, samples=2] coordinates {(1,-0.5)(1,3.5)};
    \foreach \lc in {0, ..., 3}{
      \modelboxplot{lstm-3-\lc-ig}{violet}{\lc+0.2}
      \modelboxplot{transformer-3-\lc-ig}{teal}{\lc-0.2}
    \legend{,LSTM, Transformer}
    }
  \end{axis}
  \begin{axis}[name=ppl,
    anchor=north west,at={($(ig.south west)-(0,1.1cm)$)},
    width=8cm, height=2.5cm, 
    boxplot/draw direction=x, axis x line*=bottom, hide y axis,
    xlabel=$\Ppl_{M_j}$,
    ]
    \modelboxplot{lstm-3-ppl}{violet}{0.2}
    \modelboxplot{transformer-3-ppl}{teal}{-.2}
  \end{axis}
\end{tikzpicture}
\caption{
  Average model performance by community.
The boxes indicate upper and lower quartiles, while the whiskers are placed at the upper and lower maximum, with communities more than $1.5\times IQR$ (inter-quartile range) above the upper quartile considered outliers (represented as dots). The three most extreme outliers are excluded from this view.}% \cref{fig:ppl-info-gain}).
   %See also table 1 in the supplementary materials % not very useful for this actually since we sort by indisc and it's only one model anyway.
   %for a tabular version of these results with community names.
\label{fig:comm-stratified-box}
\end{figure}

We observe empirically that across all the models we
tested, communities where conditioning has the least effect tend to be
organised around more general interest topics, such as 
\subreddit{relationships} and \subreddit{advice}, where the subject
matter is relevant to a broad range of people.  Conditioning the model
on community appears to have the most benefit for narrower special-interest
subreddits, such as those organised around a certain videogame, sports
team, or subculture.  This supports the idea that communities of practice
are the primary locus of linguistic variation.
%makes sense intuitively, since communities
%with more niche topics would tend to have more specialised vocabulary
%and other linguistic patterns.

@@

In this section we investigate the degree to which CCLM community
embeddings correlate with the community-member relationship on Reddit.

To this end, we compare the CCLM-learned community
embeddings\footnote{
  In this section, we only consider the embeddings from 
  the \emph{best} (highest information gain) CCLM from each architecture family; 
  that is, the LSTM with $l_c=1$ and the Transformer with $l_c=0$, however we
  observed similar results for other values of $l_c$.}
with the community embedding created by \citet{Kumar2018},\footnote{
  Available at \url{https://snap.stanford.edu/data/web-RedditEmbeddings.html}}
which were generated using using a negative-sampling optimization
algorithm, with the message author-community co-occurrence matrix as ground
truth, using data from January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only reflect the social
relationship between communities, via 
the user-community membership relation. In contrast, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%

@@

Unsuprisingly, the first category (\cref{fig:embedding-differences-examples}, left) yields communities that are qualitatively very similar. 
The \subreddit{SSBPM} and \subreddit{darksouls} communities are focused
around discussion of a particular videogame, and are paired
with communities that discuss a variation of the same game.
The \subreddit{amiugly} and \subreddit{Rateme} communities are both 
forums where the posts are selfies and the comments are mostly 
comments on the person's appearance.
The two communities paired with \subreddit{reddevils} are 
likewise comprised of fans of a particular English football club.
That the three football club communities are so closely related points
to an important point about the social embedding: Because of the way
it was trained, it does not have to be the case that communities with
a similar representation have a high overlap in membership,
but rather that members from the first community look similar
to members from the second community in terms of their \emph{overall}
community membership profile.

Communities with similar linguistic embeddings but dissimilar social embeddings
(\cref{fig:embedding-differences-examples}, left) tend to share a similar
topic, mode of interaction, or language variety, but 
in all cases we looked at, there is some reason to expect that
they might nevertheless attract different members.
For example,
\subreddit{hiphopheads} and \subreddit{Monstercat} are both topically
related to music, but the music genres are different, and the later
has a more geographically local focus (Monstercat is an independent
electronic music label based in Vancouver). The interactions in both 
\subreddit{MLPLounge} and \subreddit{CasualConversation} 
could be described as casual conversation, the former is intended specifically for
members of a niche internet sub-culture. % adult fans of the children's television show My Little Pony. There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
The \subreddit{exmormon} and \subreddit{Catholicism} communities discuss
the Mormon and Catholic churches, although their members have different 
relationships towards those organizations---the former is intended for
former members of the church, whereas the later is geared towards practicing
Catholics.
Finally, both \subreddit{rocketbeans} and \subreddit{de} are primarily
German-language subreddits, but the former is comprised of fans of a
computer gaming YouTube channel, while the later is more
general-interest.

@@

\subsection{Comparing embeddings: Procrustes method}
\label{sec:statistical-evidence}

The anecdotal evidence suggests that for some communities, proximity
in linguistic patterns do match social proximity and for others it
does not. Unfortunately, it seems hard, if not impossible, to draw a
general conclusion from this mode of analysis. In this section, we propose a
systematic approach with which we can quantify the correlation
between social proximity and linguistic proximity, and measure its
statistical significance.

@@

Here, the orthogonal matrix \(Ω\) gives a map from linguistic embeddings to social embeddings.
The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}.\footnote{
This approach has also been used to compare word embeddings across representations \citep[e.g.,][]{Hamilton2016a}.}
The solution is
\[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
decomposition (SVD) \(U^TΣV = LS^T\). The vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). That is, each
singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

As is common when doing SVD, we arrange \(U\), \(V\) and \(Σ\) such
that \(σ_i > σ_j\) iff \(i < j\). Doing so, the largest singular value
\(σ_0\) corresponds to the principal directions of correlation
(\(U_0\), \(V_0\)), \(σ_1\) to the second principal direction, etc.

@@

Thus if the real \(d(L,S)\) is below the mean by several standard
deviations, we can safely assume that there is statistically
significant correlation between \(L\) and \(S\). A 4-sigma difference
has less than one percent chance of occurring randomly. In our case,
we observe a difference of between 61 and 68 standard deviations
(\cref{tab:embedding-correlations}). This definitely indicates a
significant correlation.  Furthermore, by coming back to the
definition of \(d(L,S)\), we know that, on average, the cosine
similarity between \(ΩL\) and \(S\) is \(0.45 = \) (\(510 - 232.18 / 510\)).  It
further means that if we obtain a linguistic embedding \(L_k\) for a
new community $k$, we can estimate its social embedding by \(ΩL_k\),
and the cosine similarity with its true social embedding \(S_k\) is
expected to be \(0.39\) = \((431.39 - 232.18) / 510\) --- accounting for
over-fitting effects by taking the average distance rather than the
maximum. % JP: TODO is the number from the table. this should probably be rewritten
In sum, it is clear that the CCLM embeddings predict some aspect the social-network embeddings--- but far from all of it.

@@

There are several recent studies that aim to measure \emph{linguistic distinctiveness}
at the level of speech community \citep{OConnor2010,Zhang2018,Lucy2021}, 
which is an possible interpretation of the community-stratified information gain 
of the CCLM over its unconditioned counterpart \cref{sec:ig}. 
Whereas the metrics in previous work are based on lexical frequency
(and in the case of \citet{Lucy2021}, word sense distributions),
CCLM information gain is capable of capturing 
distinctiveness at multiple levels of linguistic analysis. 
However, further work is needed to
investigate exactly what kinds of variation are captured.

@@

The pattern of information gain by community is similar across
architectures;  communities that benefit most from the conditioned
model behave that way for both the LSTM and transformer.  However,
there are some differences.  For example, many of the communities with the biggest 
difference in information gain between the $l_c=0$ and $l_c=3$ LSTMs 
are organised around trading collectables or organising virtual meetups
(e.g.,
%\subreddit{friendsafari},
%\subreddit{Fireteams},
%\subreddit{RandomActsOfGaming},
%\subreddit{randomactsofcsgo},
\subreddit{Pokemongiveaway},
\subreddit{ACTrade}, and
\subreddit{SVExchange}).  
These communities tended to have highly conventionalized ways negotiating trades
and coordinating meetups.
%Why was the model where the community information was incorporated before
%the LSTM able to take better advantage of it for these communities in particular? 
It would be interesting to investigate these differences further in future work, since it
could reveal differences in the kind of linguistic variation the
different model architectures capture.

@@

Our main result is that community representations learned by
CCLMs are positively correlation with user co-occurrence patterns.
Even though such \emph{homophilic} correlation is a core hypothesis of 
sociolinguistics (see \citet{Kovacs2020}, for example), 
we believe that this study is the first to test it at 
the level of communities of practice using computational methods.
Furthermore, it appears that our method (correlating
linguistic embeddings and social embeddings) is novel. Indeed, even
though the Procrustes method has been used to correlate two sets of
linguistic embeddings \emph{for the same model}, we find no evidence
of the method being applied to embeddings for widely different models,
as we have done.

@@

\section{Ethical considerations}\label{sec:ethics}
